\documentclass[12pt]{article}

\usepackage{sbc-template}
\usepackage{graphicx,url}
\usepackage{xcolor}
\usepackage{placeins}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
\usepackage[brazil]{babel}

%\newcommand{\todo}[1]{} % Uncomment this line to hide TODO notes
\newcommand{\todo}[1]{\textcolor{red}{#1}} % TODO notes will be shown in red


\sloppy

\title{Developing and Evaluating Predictive Models for Chronic Kidney Disease Detection}

\author{Ariel M. Fagundes\inst{1}, Rodrigo\inst{1}, Anderson\inst{1} }

\address{Instituto de Informática -- Universidade Federal do Rio Grande do Sul
  (UFRGS)\\
  Caixa Postal 15.064 -- 91.501-970 -- Porto Alegre -- RS -- Brazil
  \email{\{ariel,rodrigo,anderson\}@inf.ufrgs.br}
}

\begin{document}

\maketitle

\begin{abstract}
\todo{
\begin{itemize}
    \item Briefly introduce the problem being addressed and its practical relevance.
    \item State the main objective of the study.
    \item Summarize the dataset used and the general machine learning approach adopted.
    \item Mention the main algorithms or techniques applied.
    \item Present the most important results, including the best model performance using the primary evaluation metric.
    \item Highlight the main contribution or insight obtained from the study.
\end{itemize}
}

Chronic Kidney Disease (CKD) is a relevant public health problem in which early diagnosis may
improve patient outcomes. This work evaluates machine learning models for CKD prediction using the
UCI CKD dataset \cite{risk_factor_prediction_of_chronic_kidney_disease_857}. Random Forest, SVM,
KNN, Naive Bayes, and Decision Tree were evaluated with balancing and dimensionality reduction
techniques, including SMOTE, Random Under Sampling, SMOTE + Tomek Links, PCA, SelectKBest,
SelectFromModel, and RFECV. Models were assessed using accuracy, precision, recall, and F1-score.
The best-performing models were further analyzed using explainability and interpretability techniques
to provide insights into the decision-making process and feature importance for CKD prediction.

\end{abstract}

% ===== Sections =====

\input{sections/introduction}

\input{sections/methodology}

\input{sections/results}

\input{sections/conclusion}

% ===== References =====
\section{References}
\bibliographystyle{sbc}
\bibliography{references}

\end{document}
