\section{Conclusion}

This study evaluated machine learning models for Chronic Kidney Disease
prediction through four stages: classifier comparison, preprocessing analysis,
Random Forest hyperparameter tuning, and post-hoc interpretability. The results
showed that Random Forest was the best-performing classifier among the evaluated
models, reaching overall accuracy of 0.9950 and macro-F1 of 0.9945 under
stratified cross-validation. It also achieved strong class-wise performance,
with F1-score of 0.9961 for CKD and 0.9930 for NOTCKD.

The preprocessing experiments clarified how balancing and dimensionality
reduction affected performance on this dataset. Class-balancing strategies did
not improve the selected workflow, suggesting that, under the evaluated setup,
the original class distribution did not prevent the models from achieving high
predictive performance. Dimensionality reduction had a more nuanced effect: PCA
improved the average performance in some experiments and produced the highest
mean F1-score for Random Forest, but the gain over the no-PCA Random Forest was
only 0.0009. Because this difference was small and PCA reduces the direct
interpretability of the original clinical variables, the final workflow selected
the default Random Forest without balancing and without PCA.

The hyperparameter-optimization experiments showed no meaningful improvement
from GridSearchCV or RandomizedSearchCV over the default Random Forest
configuration. The best PCA-based GridSearchCV result matched the PCA-based
baseline, while the no-PCA optimized configurations remained equivalent to the
no-PCA baseline.

This configuration kept predictive performance nearly unchanged while
preserving a simpler workflow and explanations based on the original attributes.
The interpretability analysis highlighted hematological, renal-function,
urine-related, and clinical variables as relevant to model predictions,
including hemo, pcv, sg, grf, al, rbcc, htn, dm, and bgr. These findings are
consistent with clinically relevant CKD markers, but they should be interpreted
as model-dependent predictive patterns rather than causal evidence.

The main limitation of this study is the small dataset size, with 199 instances
after cleaning. This restricts the ability to generalize the results to broader
clinical populations and makes external validation especially important. In
addition, the dataset contains range-encoded attributes rather than raw
continuous measurements, which may limit the granularity of the learned
patterns.

Future work should evaluate the workflow on larger and more diverse CKD
datasets and include external validation to assess generalization beyond the
dataset used in this study. Additional experiments should investigate
alternative preprocessing orders, especially applying dimensionality reduction
before class-balancing techniques. This would make it possible to compare
pipelines such as scaling followed by PCA and then SMOTE against pipelines that
apply balancing before dimensionality reduction. These alternative orders should
be evaluated within the cross-validation pipeline to avoid potential
information leakage between training and validation folds. Further work should
also consider different PCA variance thresholds, calibration and
threshold-selection strategies, confidence intervals or statistical tests for
small metric differences, and clinical review of the most relevant features
identified by the interpretability methods.
